---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# rodm2

The goal of **rodm2** is to make it easy to use the [ODM2](https://github.com/ODM2/ODM2) data model in R, in order to promote collaboration between ecologists, hydrologists, and soil scientists. 

Organize  site, sample, and sensor data all in one place! For example:

* time series hydrologic observations from a well or stream gage
* sediment sample properties from a soil core analyzed in a lab
* categorical observations manually recorded in the field

This project is just getting started! Get in touch or open an [issue](https://github.com/khondula/rodm2/issues) if you have a use case or if you're interested in collaborating. 

## Installation

You can install this package from GitHub with:

```{r, message=FALSE}
# install.packages("devtools")
devtools::install_github("khondula/rodm2")
```

## Usage

`rodm2` is designed to work either with an existing ODM2 database on a server (e.g. PostgreSQL), or with spreadsheet files that aren't (yet!) in a formal database. 

Initialize an empty sqlite database with odm2 schema in your working directory (empty except for the controlled vocabulary tables). You can view and edit this database using [DB Browser for SQLite](https://sqlitebrowser.org/) outside of R.

```{r, eval=FALSE}
library(rodm2)
create_sqlite(dir = ".")
```

To work with it in R, create a connection object

```{r}
dblite <- DBI::dbConnect(RSQLite::SQLite(), "odm2.sqlite")
head(RSQLite::dbListTables(dblite))
```
Add information using `db_describe_%TABLE%` functions
```{r}
rodm2::db_describe_person(db = dblite, PersonFirstName = "Wendy",
     PersonLastName = "Wetland",
     AffiliationStartDate = "2018-01-01",
     PrimaryEmail = "wendy 'at' swamps.edu")
```


## Acknowledgements

This project is made possible through support from [rOpenSci](https://ropensci.org/) and [SESYNC](https://www.sesync.org/). 

Read about the details and development of ODM2 in the open access paper in Environmental Modelling & Software: 

> Horsburgh, J. S., Aufdenkampe, A. K., Mayorga, E., Lehnert, K. A., Hsu, L., Song, L., Spackman Jones, A., Damiano, S. G., Tarboton, D. G., Valentine, D., Zaslavsky, I., Whitenack, T. (2016). [Observations Data Model 2: A community information model for spatially discrete Earth observations](http://dx.doi.org/10.1016/j.envsoft.2016.01.010), Environmental Modelling & Software, 79, 55-74, http://dx.doi.org/10.1016/j.envsoft.2016.01.010

