---
title: "Add new sites"
author: "Kelly Hondula"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Add new sites}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In ODM2 lingo, sites are a type of sampling feature. They must have a unique UUID, short code name (max 50 characters), and be classified as one of the pre-defined [types of sampling features](http://vocabulary.odm2.org/samplingfeaturetype/) such as "Site" or "Weather station". They can also have a longer name (max 255 characters), even longer description (max 5000 characters), and spatial/geometry and elevation information. 

All actions must be performed at/on/to a sampling feature! 

## Add a site without any geographic info

## Add points from lat lon data

```{r}
# db_insert_sites_latlon(new_sites)
```

## Add points from a kmz (Google Earth) file

## Add relationships between sites

If sites are related to each other, make use of the `relatedfeatures` table by specifying parent-child or other types of relationships. 

This allows for making a network graph showing how sites are related to each other.  

## Add Site Group annotations

Add annotations about groups of sites. 
